---
title: "Sleeping Patterns"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include = FALSE, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
library(tidyverse)

```

This data set includes four different measures of sleep: (descriptions taken directly from the CMU S&DS Data Repository)

* `bedtime_mssd`:	Mean successive squared difference of bedtime, a measure of variability in bedtimes between consecutive nights
* `total_sleep_time`:	Average time in bed (the difference between wake time and bedtime) minus the length of total awake/restlessness in the main sleep episode, in minutes
* `midpoint_sleep`:	Average midpoint of bedtime and wake time, in minutes after 11 pm (for example, 364 is 5:04 am)
* `daytime_sleep`:	Average sleep time outside of the range of the main sleep episode, including short naps or sleep that occurred during the daytime, in minutes

Below is a 

```{r}

#View(sleep_df)
                          

```


## bedtime_mssd

```{r}

#sleep_df %>% ggplot(aes(x = bedtime_mssd, fill = university)) + geom_histogram()


```

# total_sleep_time

```{r}

#sleep_df %>% ggplot(aes(x = total_sleep_time, fill = university)) + geom_density(alpha = 0.5)

```

# midpoint_sleep 

```{r}

```

# daytime_sleep

```{r}


```

