---
title: "Statistic Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(plotly)
library(ggplot2)
library(broom)
library(gtsummary)
library(knitr)
library(patchwork)
library(kableExtra)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

In this page, we use some typical statistic models to further explore the impact of students' sleeping patterns on their GPA. 

---

## Simple Linear Regression {.tabset}

To begin with, fit simple linear regression model to demonstrate the individual impact of each sleeping measure on students' GPA. Each time we use one sleeping measure as the predictor to predict the `term_gpa`.

Below are plots and statistics of the simple linear regression models. 


```{r, include=FALSE}
sleep_df <- read_csv("data/cmu-sleep.csv") %>% 
  janitor::clean_names() %>%  
  mutate(demo_race = case_when(demo_race == 0 ~ "Underrepresented", 
                                     demo_race == 1 ~ "Non-underrepresented"),
         demo_race = fct_relevel(demo_race, "Underrepresented"),
         demo_gender = case_when(demo_gender == 0 ~ "Male", 
                                     demo_gender == 1 ~ "Female"),
         demo_gender = fct_relevel(demo_gender, "Male"), 
         demo_firstgen = case_when(demo_firstgen == "0" ~ "Non-first gen", 
                                   demo_firstgen == "1" ~ "First-gen", 
                                   TRUE ~ NA),
         demo_firstgen = fct_relevel(demo_firstgen, "Non-first gen"), 
         time_collection = case_when(cohort == "lac1" ~ "Spring, 2018", 
                            cohort == "lac2" ~ "Spring, 2017",
                            cohort == "nh" ~ "Spring, 2016",
                            cohort == "uw1" ~ "Spring, 2018", 
                            cohort == "uw2" ~ "Spring, 2019"), 
         university = case_when(cohort == "lac1" ~ "CMU", 
                            cohort == "lac2" ~ "CMU",
                            cohort == "nh" ~ "NDU",
                            cohort == "uw1" ~ "UW", 
                            cohort == "uw2" ~ "UW")) |> 
  mutate(total_sleep_time = total_sleep_time / 60,
         bedtime_mssd = log(bedtime_mssd+1),
         midpoint_sleep = midpoint_sleep / 60,
         daytime_sleep =  daytime_sleep / 60)

```



### Total Sleep Time {.tabset}

**Numerical Result**

```{r}
record_tab = data.frame(c())
lm_sp = lm(term_gpa ~ total_sleep_time, data = sleep_df)

tab_sp = lm_sp |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high, -statistic) 


tab_sp|> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "3cm")

re_new = data.frame(`Single Predictor` = 'Total Sleep Time',
                    `P-value` = tab_sp$p.value[2],
                    MSE = mean(residuals(lm_sp)^2)
                    )
record_tab = rbind(record_tab, re_new)
```

**Model Visualization**

```{r}
plot_sp = 
  ggplot(data = sleep_df, aes(x = total_sleep_time, y = term_gpa)) +
  geom_point(color = "#664785", alpha = 0.5) +                      
  geom_smooth(method = "lm", se = TRUE, color = '#F2A354') +  
   geom_text(
    aes(label = paste0("MSE = ", round(mean(residuals(lm_sp)^2),6))),
    x = 8, y = 1,  
    hjust = 0, vjust = 0, size = 4, color = "#664785") +
  labs(  
    x = "Total Sleep Time (h)",
    y = "Term GPA",
    title = "Simple Linear Regression"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(plot_sp)
```




--- 

### Variability in Bed Times {.tabset}

**Numerical Result**

```{r}
lm_sp = lm(term_gpa ~ bedtime_mssd, data = sleep_df)
tab_sp = lm_sp |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high, -statistic) 


tab_sp|> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "3cm")

re_new = data.frame(`Single Predictor` = 'Variability in Bed Times',
                    `P-value` = tab_sp$p.value[2],
                    MSE = mean(residuals(lm_sp)^2)
                    )
record_tab = rbind(record_tab, re_new)
```

**Model Visualization**

```{r}
plot_sp = 
  ggplot(data = sleep_df, aes(x = bedtime_mssd, y = term_gpa)) +
  geom_point(color = "#664785", alpha = 0.5) +                      
  geom_smooth(method = "lm", se = TRUE, color = '#F2A354') +  
  geom_text(
   aes(label = paste0("MSE = ", round(mean(residuals(lm_sp)^2),6))),
   x = 2.2, y = 1,  
   hjust = 0, vjust = 0, size = 4, color = "#664785") +  
  labs(  
    x = "log(Mean Successive Squared Difference + 1)",
    y = "Term GPA",
    title = "Simple Linear Regression"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(plot_sp)
```


--- 


### Sleeping Midpoint {.tabset}

**Numerical Result**

```{r}
lm_sp = lm(term_gpa ~ midpoint_sleep, data = sleep_df)

tab_sp = lm_sp |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high, -statistic) 


tab_sp|> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "3cm")

re_new = data.frame(`Single Predictor` = 'Midpoint Sleep',
                    `P-value` = tab_sp$p.value[2],
                    MSE = mean(residuals(lm_sp)^2)
                    )
record_tab = rbind(record_tab, re_new)


```

**Model Visualization**

```{r}
plot_sp = 
  ggplot(data = sleep_df, aes(x = midpoint_sleep, y = term_gpa)) +
  geom_point(color = "#664785", alpha = 0.5) +                      
  geom_smooth(method = "lm", se = TRUE, color = '#F2A354') +  
  geom_text(
   aes(label = paste0("MSE = ", round(mean(residuals(lm_sp)^2),6))),
   x = 10, y = 1,  
   hjust = 0, vjust = 0, size = 4, color = "#664785") +  
  labs(  
    x = "Midpoint Sleep (hrs past 11 PM)",
    y = "Term GPA",
    title = "Simple Linear Regression"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(plot_sp)
```



--- 


### Daytime Sleep {.tabset}

**Numerical Result**

```{r}
lm_sp = lm(term_gpa ~ daytime_sleep, data = sleep_df)

tab_sp = lm_sp |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high, -statistic) 


tab_sp|> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "3cm")

re_new = data.frame(`Single Predictor` = 'Daytime Sleep',
                    `P-value` = tab_sp$p.value[2],
                    MSE = mean(residuals(lm_sp)^2)
                    )
record_tab = rbind(record_tab, re_new)

```


**Model Visualization**

```{r}
plot_sp = 
  ggplot(data = sleep_df, aes(x = daytime_sleep, y = term_gpa)) +
  geom_point(color = "#664785", alpha = 0.5) +                      
  geom_smooth(method = "lm", se = TRUE, color = '#F2A354') +  
  geom_text(
   aes(label = paste0("MSE = ", round(mean(residuals(lm_sp)^2),6))),
   x = 3.5, y = 1,  
   hjust = 0, vjust = 0, size = 4, color = "#664785") +  
  labs(  
    x = "Daytime Sleep (h)",
    y = "Term GPA",
    title = "Simple Linear Regression"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(plot_sp)

```

---

 **COMMENTS:**
 
We collect the p-value and the Mean Squared Error from each simple linear regression model to the following table:


```{r}
record_tab|> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "5cm")
```

Sleeping measure 'Total Sleep Time', 'Midpoint Sleep', and 'Daytime Sleep' have p-value less than 0.05, which means they have coefficient significantly not equal to 0. However, the measure 'Variability in Bed Times' has p-value larger than 0.05, implies this variable doesn't have significant impact on `term_gpa`.



---

## Multiple Linear Regression

Now consider including more variables to fit linear regression

### Sleeping Measures Only{.tabset}

Here use all the 4 sleeping measures as predictors to predict `term_gpa`.

```{r}
set.seed(100)
lm_fit = lm(term_gpa ~ bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep, data = sleep_df)

lm_result =
  broom::tidy(lm_fit,conf.int = TRUE) |> 
  select(term, estimate, std.error, conf.low, conf.high) |> 
  mutate(Method = 'Linear Regression')|> 
  mutate(term = reorder(term, estimate))

boot_straps = 
  sleep_df|> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result = 
  boot_straps |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(Method = 'Bootstrap',
         estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high,Method)|> 
  mutate(term = reorder(term, estimate))
  

com_coef =
  rbind(lm_result,boot_result) |> 
  filter(term != "(Intercept)") 

plot_coef = 
  com_coef |>   
  ggplot(aes(x = term, estimate, y = estimate, color = Method)) +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.5, 
                position = position_dodge(width = 0.3)) +
  coord_flip() +
  labs(
    title = "Effect Sizes of Predictors on Cumulative GPA",
    x = "Predictors",
    y = "Coefficient Estimate"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

plot_coef

```

**Comments**

For Sleep Measure Only  

The detailed information of the estimated coefficient are listed below:


#### Linear Regression{.tabset}

```{r}
lm_result |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  arrange(-estimate) |> 
  select(-Method) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4cm") 

```

---

#### Bootstrap{.tabset}

```{r}
boot_result |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-Method) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4cm") 
```


---

### Students' Characteristics and Sleeping Measures{.tabset}


```{r,fig.width= 6, fig.height=10 }

set.seed(100)
lm_fit = lm(term_gpa ~ university + demo_race + demo_gender + demo_firstgen + bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep + cum_gpa, data = sleep_df)

lm_result =
  broom::tidy(lm_fit,conf.int = TRUE) |> 
  select(term, estimate, std.error, conf.low, conf.high) |> 
  mutate(Method = 'Linear Regression')|> 
  mutate(term = reorder(term, estimate))

boot_straps = 
  sleep_df|> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ university + demo_race + demo_gender + demo_firstgen + bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep + cum_gpa, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result = 
  boot_straps |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(Method = 'Bootstrap',
         estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high,Method) |> 
  mutate(term = reorder(term, estimate))
  

com_coef =
  rbind(lm_result,boot_result) |> 
  filter(term != "(Intercept)") 

plot_coef = 
  com_coef |>   
  ggplot(aes(x = term, y = estimate, color = Method)) +
  geom_point(position = position_dodge(width = 0.7)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.5, 
                position = position_dodge(width = 0.7)) +
  coord_flip() +
  labs(
    title = "Effect Sizes of Predictors on Cumulative GPA",
    x = "Predictors",
    y = "Coefficient Estimate"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

plot_coef


```

**Comments**

According to the regression result, the variable `cum_gpa`, `universityNDU`, `total_sleep_time` are significant non-zero, which means it has great impact on `term_gpa`. It make sense since the `cum_gpa` reflects the students academic background before the studied term, which is crucial to the students' performance in a new term. Additionally, different universities have different academic atmosphere and also follows different grading patterns for students' study performance, leading to distinct `term_gpa` even when other variables are the same. Besides, the larger `total_sleep_time` have potential to improve students' study performance, which implies getting a good sleep is helpful in increasing the GPA.

The detailed information of the estimated coefficient are listed below:


#### Linear Regression{.tabset}

```{r}
lm_result |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-Method) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

```

---

#### Bootstrap{.tabset}

```{r}
boot_result |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-Method) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 
```

---


## University Specific Model


From the multiple linear regression result, the `university` variable has significant impact on the `term_gpa`. Thus, we're going to further explore the how the other predictors impact the `term_gpa` within a specific university.


### Sleeping Measures Impact{.tabset}

The following figures shows the individual impact of each sleeping measure within different universities. The each line in the plot represents a simple linear regression model for a specific university.  


#### Total Sleep Time {.tabset}

```{r,fig.width= 6, fig.height=10}
plot_uni = 
  ggplot(sleep_df, aes(x = total_sleep_time, y = term_gpa)) +
  geom_point(alpha = 0.6, color = "#664785") +
  geom_smooth(method = "lm", se = FALSE, aes(color = university),alpha = 0.5) +
  labs(
    title = "University Specific Simple Linear Regression",
    x = "Total Sleep Time (Hours)",
    y = "Term GPA"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
ggplotly(plot_uni)

```



--- 

#### Variability in Bed Times {.tabset}


```{r}
plot_uni = 
  ggplot(sleep_df, aes(x = bedtime_mssd, y = term_gpa)) +
  geom_point(alpha = 0.6, color = "#664785") +
  geom_smooth(method = "lm", se = FALSE, aes(color = university),alpha = 0.5) +
  labs(
    title = "University Specific Simple Linear Regression",
    x = "Mean Successive Squared Difference",
    y = "Term GPA"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
ggplotly(plot_uni)
```


--- 


#### Sleeping Midpoint {.tabset}


```{r}
plot_uni = 
  ggplot(sleep_df, aes(x = midpoint_sleep, y = term_gpa)) +
  geom_point(alpha = 0.6, color = "#664785") +
  geom_smooth(method = "lm", se = FALSE, aes(color = university),alpha = 0.5) +
  labs(
    title = "University Specific Simple Linear Regression",
    x = "Midpoint Sleep (hrs past 11 PM)",
    y = "Term GPA"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
ggplotly(plot_uni)
```


--- 


#### Daytime Sleep {.tabset}


```{r}
plot_uni = 
  ggplot(sleep_df, aes(x = daytime_sleep, y = term_gpa)) +
  geom_point(alpha = 0.6, color = "#664785") +
  geom_smooth(method = "lm", se = FALSE, aes(color = university),alpha = 0.5) +
  labs(
    title = "University Specific Simple Linear Regression",
    x = "Daytime Sleep (h)",
    y = "Term GPA"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
ggplotly(plot_uni)
```

---

 **COMMENTS:**
 
## Model Fitting

Now we fit the multiple linear regression model for each university and we fit the linear regression model based on 1000 times of bootstrap.

### Sleep Measure Only{.tabset}

When the predictors include only the 4 sleeping measure variables, we have the following results



```{r}
set.seed(100)

boot_straps_cmu = 
  sleep_df|> 
  filter(university == "CMU") |> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result_cmu = 
  boot_straps_cmu |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(Method = 'Bootstrap',
         estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high,Method) |> 
  mutate(term = reorder(term, estimate),
         university = 'CMU')


boot_straps_ndu = 
  sleep_df|> 
  filter(university == "NDU") |> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result_ndu = 
  boot_straps_ndu |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(Method = 'Bootstrap',
         estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high,Method) |> 
  mutate(term = reorder(term, estimate),
         university = 'NDU')


boot_straps_uw = 
  sleep_df|> 
  filter(university == "UW") |> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result_uw = 
  boot_straps_uw |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(Method = 'Bootstrap',
         estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high,Method) |> 
  mutate(term = reorder(term, estimate),
         university = 'UW')


uni_coef = rbind(boot_result_cmu,
                 boot_result_ndu,
                 boot_result_uw) |> 
  mutate(university = factor(university, levels = c("CMU", "NDU", "UW"))) 

uni_coef |>
  filter(term != "(Intercept)") |> 
  ggplot(aes(x = term, y = estimate, color = university)) +
  geom_point(position = position_dodge(width = 0.7)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.5, 
                position = position_dodge(width = 0.7)) +
  coord_flip() +
  labs(
    title = "Effect Sizes of Predictors on Cumulative GPA",
    x = "Predictors",
    y = "Coefficient Estimate"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

```

As we can see from the coefficient estimation result, there's different pattern of how sleeping measures affect the students' term GPA. And more detailed result are 


#### CMU{.tabset}


```{r}
boot_result_cmu |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-university) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

```



#### NDU{.tabset}


```{r}
boot_result_ndu |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-university) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

```


---



#### UW{.tabset}


```{r}
boot_result_uw |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-university) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

```

---
