---
title: "Statistic Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(plotly)
library(ggplot2)
library(broom)
library(gtsummary)
library(knitr)
library(patchwork)
library(kableExtra)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

In this page, we further explore the impact of students' sleeping patterns on their GPA by fitting GPA prediction models based on the predictors we're interested. We consider using linear regression model combined with bootstrapping method for model fitting. 

---

## 1. Simple Linear Regression {.tabset}

To begin with, fit simple linear regression model to demonstrate the individual effect of each sleeping measure on students' term GPA. Each time we use one sleeping measure as the predictor to predict the `term_gpa`.

The result shows that the `total_sleep_time` (see tab *Total Sleep Time*) variable has positive correlation with `term gpa` while the other three variables -- `bedtime_mssd`(see tab *Variability in Bed Times*), `midpoint_sleep`(see tab *Sleep Midpoint*) and `daytime_sleep`(see tab *Daytime Sleep*) are negative related with `term_gpa`. We also collect the p-value and the Mean Squared Error (MSE) from each simple linear regression model(see tab *Comparison*). All of the sleeping measures have significantly non-zero coefficient when used as the only predictor to predict term GPA. Among them, the simple linear regression has the smallest MSE when using the variable `total_sleep_time`.

Below are detailed information about the simple linear regression models.

```{r, include=FALSE}
sleep_df <- read_csv("data/cmu-sleep.csv") %>% 
  janitor::clean_names() %>%  
  mutate(demo_race = case_when(demo_race == 0 ~ "Underrepresented", 
                                     demo_race == 1 ~ "Non-underrepresented"),
         demo_race = fct_relevel(demo_race, "Underrepresented"),
         demo_gender = case_when(demo_gender == 0 ~ "Male", 
                                     demo_gender == 1 ~ "Female"),
         demo_gender = fct_relevel(demo_gender, "Male"), 
         demo_firstgen = case_when(demo_firstgen == "0" ~ "Non-first gen", 
                                   demo_firstgen == "1" ~ "First-gen", 
                                   TRUE ~ NA),
         demo_firstgen = fct_relevel(demo_firstgen, "Non-first gen"), 
         time_collection = case_when(cohort == "lac1" ~ "Spring, 2018", 
                            cohort == "lac2" ~ "Spring, 2017",
                            cohort == "nh" ~ "Spring, 2016",
                            cohort == "uw1" ~ "Spring, 2018", 
                            cohort == "uw2" ~ "Spring, 2019"), 
         university = case_when(cohort == "lac1" ~ "CMU", 
                            cohort == "lac2" ~ "CMU",
                            cohort == "nh" ~ "NDU",
                            cohort == "uw1" ~ "UW", 
                            cohort == "uw2" ~ "UW")) |> 
  mutate(total_sleep_time = total_sleep_time / 60,
         bedtime_mssd = log(bedtime_mssd+1),
         midpoint_sleep = midpoint_sleep / 60,
         daytime_sleep =  daytime_sleep / 60)

```



### Total Sleep Time {.tabset}

**Numerical Result**

```{r}
record_tab = data.frame(c())
lm_sp = lm(term_gpa ~ total_sleep_time, data = sleep_df)

tab_sp = lm_sp |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high, -statistic) 


tab_sp|> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "3cm")

re_new = data.frame(`Single Predictor` = 'Total Sleep Time',
                    `P-value` = tab_sp$p.value[2],
                    MSE = mean(residuals(lm_sp)^2)
                    )
record_tab = rbind(record_tab, re_new)
```

**Model Visualization**

```{r}
plot_sp = 
  ggplot(data = sleep_df, aes(x = total_sleep_time, y = term_gpa)) +
  geom_point(color = "#664785", alpha = 0.5) +                      
  geom_smooth(method = "lm", se = TRUE, color = '#F2A354') +  
   geom_text(
    aes(label = paste0("MSE = ", round(mean(residuals(lm_sp)^2),6))),
    x = 8, y = 1,  
    hjust = 0, vjust = 0, size = 4, color = "#664785") +
  labs(  
    x = "Total Sleep Time (h)",
    y = "Term GPA",
    title = "Simple Linear Regression"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(plot_sp)
```




--- 

### Variability in Bed Times {.tabset}

**Numerical Result**

```{r}
lm_sp = lm(term_gpa ~ bedtime_mssd, data = sleep_df)
tab_sp = lm_sp |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high, -statistic) 


tab_sp|> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "3cm")

re_new = data.frame(`Single Predictor` = 'Variability in Bed Times',
                    `P-value` = tab_sp$p.value[2],
                    MSE = mean(residuals(lm_sp)^2)
                    )
record_tab = rbind(record_tab, re_new)
```

**Model Visualization**

```{r}
plot_sp = 
  ggplot(data = sleep_df, aes(x = bedtime_mssd, y = term_gpa)) +
  geom_point(color = "#664785", alpha = 0.5) +                      
  geom_smooth(method = "lm", se = TRUE, color = '#F2A354') +  
  geom_text(
   aes(label = paste0("MSE = ", round(mean(residuals(lm_sp)^2),6))),
   x = 2.2, y = 1,  
   hjust = 0, vjust = 0, size = 4, color = "#664785") +  
  labs(  
    x = "log(Mean Successive Squared Difference + 1)",
    y = "Term GPA",
    title = "Simple Linear Regression"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(plot_sp)
```


--- 


### Sleeping Midpoint {.tabset}

**Numerical Result**

```{r}
lm_sp = lm(term_gpa ~ midpoint_sleep, data = sleep_df)

tab_sp = lm_sp |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high, -statistic) 


tab_sp|> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "3cm")

re_new = data.frame(`Single Predictor` = 'Midpoint Sleep',
                    `P-value` = tab_sp$p.value[2],
                    MSE = mean(residuals(lm_sp)^2)
                    )
record_tab = rbind(record_tab, re_new)


```

**Model Visualization**

```{r}
plot_sp = 
  ggplot(data = sleep_df, aes(x = midpoint_sleep, y = term_gpa)) +
  geom_point(color = "#664785", alpha = 0.5) +                      
  geom_smooth(method = "lm", se = TRUE, color = '#F2A354') +  
  geom_text(
   aes(label = paste0("MSE = ", round(mean(residuals(lm_sp)^2),6))),
   x = 10, y = 1,  
   hjust = 0, vjust = 0, size = 4, color = "#664785") +  
  labs(  
    x = "Midpoint Sleep (hrs past 11 PM)",
    y = "Term GPA",
    title = "Simple Linear Regression"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(plot_sp)
```



--- 


### Daytime Sleep {.tabset}

**Numerical Result**

```{r}
lm_sp = lm(term_gpa ~ daytime_sleep, data = sleep_df)

tab_sp = lm_sp |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high, -statistic) 


tab_sp|> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "3cm")

re_new = data.frame(`Single Predictor` = 'Daytime Sleep',
                    `P-value` = tab_sp$p.value[2],
                    MSE = mean(residuals(lm_sp)^2)
                    )
record_tab = rbind(record_tab, re_new)

```


**Model Visualization**

```{r}
plot_sp = 
  ggplot(data = sleep_df, aes(x = daytime_sleep, y = term_gpa)) +
  geom_point(color = "#664785", alpha = 0.5) +                      
  geom_smooth(method = "lm", se = TRUE, color = '#F2A354') +  
  geom_text(
   aes(label = paste0("MSE = ", round(mean(residuals(lm_sp)^2),6))),
   x = 3.5, y = 1,  
   hjust = 0, vjust = 0, size = 4, color = "#664785") +  
  labs(  
    x = "Daytime Sleep (h)",
    y = "Term GPA",
    title = "Simple Linear Regression"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(plot_sp)

```

---

### Comparision{.tabset}


```{r}
record_tab|> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "5cm")
```



---







## 2. Multiple Linear Regression

In this section, we aim to leverage more predictor at a time for model fitting. The first part (**Sleeping Measures Only Model**) focus on linear model that includes all of the 4 sleeping measures as predictors without considering other factors that may have impact on term GPA. The second part (**Characteristic-Included Model**) is about the linear model that includes other informative characteristics of students as the predictors. For each part, we applied linear regression on the initial data set (labeled as *Initial*) and 1000 times bootstrapped data set (labeled as *Bootstrap*). 



### Sleeping Measures Only Model{.tabset}

```{r}
set.seed(100)
lm_fit = lm(term_gpa ~ bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep, data = sleep_df)

lm_result =
  broom::tidy(lm_fit,conf.int = TRUE) |> 
  select(term, estimate, std.error, conf.low, conf.high) |> 
  mutate(Method = 'Linear Regression')|> 
  mutate(term = reorder(term, estimate))

boot_straps = 
  sleep_df|> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result = 
  boot_straps |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(Method = 'Bootstrap',
         estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high,Method)|> 
  mutate(term = reorder(term, estimate))
  

com_coef =
  rbind(lm_result,boot_result) |> 
  filter(term != "(Intercept)") 

plot_coef = 
  com_coef |>   
  ggplot(aes(x = term, estimate, y = estimate, color = Method)) +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.5, 
                position = position_dodge(width = 0.3)) +
  coord_flip() +
  labs(
    title = "Effect Sizes of Predictors on Cumulative GPA",
    x = "Predictors",
    y = "Coefficient Estimate"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

plot_coef

```


When including only 4 sleeping measure variables as the predictors, the linear regression model based either initial data or the bootstrapped data have significant variables `total_sleep_time` and `midpoint_sleep`. However, for the model based on initial data, the coefficient of `daytime_sleep` is also significant non-zero. The detailed information of the estimated coefficient are listed below:

#### Linear Regression{.tabset}

```{r}
lm_result |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  arrange(-estimate) |> 
  select(-Method) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4cm") 

```

---

#### Bootstrap{.tabset}

```{r}
boot_result |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-Method) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4cm") 
```

---

### Characteristic-Included Model{.tabset}


```{r}

set.seed(100)
lm_fit = lm(term_gpa ~ university + demo_race + demo_gender + demo_firstgen + bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep + cum_gpa, data = sleep_df)

lm_result =
  broom::tidy(lm_fit,conf.int = TRUE) |> 
  select(term, estimate, std.error, conf.low, conf.high) |> 
  mutate(Method = 'Linear Regression')|> 
  mutate(term = reorder(term, estimate))

boot_straps = 
  sleep_df|> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ university + demo_race + demo_gender + demo_firstgen + bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep + cum_gpa, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result = 
  boot_straps |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(Method = 'Bootstrap',
         estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high,Method) |> 
  mutate(term = reorder(term, estimate))
  

com_coef =
  rbind(lm_result,boot_result) |> 
  filter(term != "(Intercept)") 

plot_coef = 
  com_coef |>   
  ggplot(aes(x = term, y = estimate, color = Method)) +
  geom_point(position = position_dodge(width = 0.7)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.5, 
                position = position_dodge(width = 0.7)) +
  coord_flip() +
  labs(
    title = "Effect Sizes of Predictors on Cumulative GPA",
    x = "Predictors",
    y = "Coefficient Estimate"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

plot_coef


```

According to the regression result, the variable `cum_gpa`, `universityNDU`, `total_sleep_time` are significantly non-zero, which means it has great impact on `term_gpa`. It makes sense since the `cum_gpa` reflects the students academic background before the studied term, which is crucial to the students' performance in a new term. Additionally, different universities have different academic atmosphere and also follows different grading patterns for students' study performance, leading to distinct `term_gpa` even when other variables are the same. Besides, the larger `total_sleep_time` have potential to improve students' study performance, which implies getting a good sleep is helpful in increasing the GPA. The detailed information of the estimated coefficient are listed below:

#### Linear Regression{.tabset}

```{r}
lm_result |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-Method) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

```

---

#### Bootstrap{.tabset}

```{r}
boot_result |> 
  filter(term != "(Intercept)") |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-Method) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 
```

---


## 3. Further Exploration

From the multiple linear regression result, the `university` variable has significant impact on the `term_gpa`. Thus, we're going to further explore the how the other predictors impact the `term_gpa` within a specific university.


### Sleeping Measures Impact{.tabset}

The following figures show the individual impact of each sleeping measure within different universities. Each line in the plot represents a simple linear regression model for a specific university.  

#### Total Sleep Time {.tabset}

```{r}
plot_uni = 
  ggplot(sleep_df, aes(x = total_sleep_time, y = term_gpa)) +
  geom_point(alpha = 0.6, color = "#664785") +
  geom_smooth(method = "lm", se = FALSE, aes(color = university),alpha = 0.5) +
  labs(
    title = "University Specific Simple Linear Regression",
    x = "Total Sleep Time (Hours)",
    y = "Term GPA"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
ggplotly(plot_uni)

```

--- 

#### Variability in Bed Times {.tabset}


```{r}
plot_uni = 
  ggplot(sleep_df, aes(x = bedtime_mssd, y = term_gpa)) +
  geom_point(alpha = 0.6, color = "#664785") +
  geom_smooth(method = "lm", se = FALSE, aes(color = university),alpha = 0.5) +
  labs(
    title = "University Specific Simple Linear Regression",
    x = "Mean Successive Squared Difference",
    y = "Term GPA"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
ggplotly(plot_uni)
```


--- 


#### Sleeping Midpoint {.tabset}


```{r}
plot_uni = 
  ggplot(sleep_df, aes(x = midpoint_sleep, y = term_gpa)) +
  geom_point(alpha = 0.6, color = "#664785") +
  geom_smooth(method = "lm", se = FALSE, aes(color = university),alpha = 0.5) +
  labs(
    title = "University Specific Simple Linear Regression",
    x = "Midpoint Sleep (hrs past 11 PM)",
    y = "Term GPA"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
ggplotly(plot_uni)
```


--- 


#### Daytime Sleep {.tabset}


```{r}
plot_uni = 
  ggplot(sleep_df, aes(x = daytime_sleep, y = term_gpa)) +
  geom_point(alpha = 0.6, color = "#664785") +
  geom_smooth(method = "lm", se = FALSE, aes(color = university),alpha = 0.5) +
  labs(
    title = "University Specific Simple Linear Regression",
    x = "Daytime Sleep (h)",
    y = "Term GPA"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
ggplotly(plot_uni)
```



### Sleep Measure Only Model{.tabset}

Now we fit the multiple linear regression model for each university and we also consider both the sleeping measure only model and characteristic-included model. Besides, since we need separate the data into three sub dataset by university for model fitting, the number of the sample size decreases. To ensure good performance of linear model, we only consider the model fitting with bootstrapping. For each university, we bootstrap the corresponding sub data for 1000 times and apply the linear regression. 

```{r}
set.seed(100)

boot_straps_cmu = 
  sleep_df|> 
  filter(university == "CMU") |> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result_cmu = 
  boot_straps_cmu |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high) |> 
  mutate(term = reorder(term, estimate),
         university = 'CMU')|>
  filter(term != "(Intercept)") 


boot_straps_ndu = 
  sleep_df|> 
  filter(university == "NDU") |> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result_ndu = 
  boot_straps_ndu |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high) |> 
  mutate(term = reorder(term, estimate),
         university = 'NDU')|>
  filter(term != "(Intercept)") 


boot_straps_uw = 
  sleep_df|> 
  filter(university == "UW") |> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result_uw = 
  boot_straps_uw |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high) |> 
  mutate(term = reorder(term, estimate),
         university = 'UW')|>
  filter(term != "(Intercept)") 


uni_coef = rbind(boot_result_cmu,
                 boot_result_ndu,
                 boot_result_uw) |> 
  mutate(university = factor(university, levels = c("CMU", "NDU", "UW"))) 

uni_coef |> 
  ggplot(aes(x = term, y = estimate, color = university)) +
  geom_point(position = position_dodge(width = 0.7)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.5, 
                position = position_dodge(width = 0.7)) +
  coord_flip() +
  labs(
    title = "Effect Sizes of Predictors on Cumulative GPA",
    x = "Predictors",
    y = "Coefficient Estimate"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

```

As we can see from the coefficient estimation result, there's different patterns of how sleeping measures affect the students' term GPA. `total_sleep_time` is significant for both NDU and UW. More detailed result are shown below:

#### CMU{.tabset}


```{r}
boot_result_cmu |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-university) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

term_order <- boot_result_cmu |> arrange(-estimate) 
term_order = term_order$term
```



#### NDU{.tabset}


```{r}
boot_result_ndu |> 
  mutate(
    term = factor(term, levels = term_order),
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-university) |> 
  arrange(term) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

```


---



#### UW{.tabset}


```{r}
boot_result_uw |> 
  mutate(
    term = factor(term, levels = term_order),
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-university) |> 
  arrange(term) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

```

---





### Characteristic-Included Model{.tabset}

```{r}
set.seed(100)

boot_straps_cmu = 
  sleep_df|> 
  filter(university == "CMU") |> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ demo_race + demo_gender + demo_firstgen + bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep + cum_gpa, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result_cmu = 
  boot_straps_cmu |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high) |> 
  mutate(term = reorder(term, estimate),
         university = 'CMU')|> 
  filter(term != "(Intercept)")


boot_straps_ndu = 
  sleep_df|> 
  filter(university == "NDU") |> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ demo_race + demo_gender + demo_firstgen + bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep + cum_gpa, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result_ndu = 
  boot_straps_ndu |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high) |> 
  mutate(term = reorder(term, estimate),
         university = 'NDU')|> 
  filter(term != "(Intercept)")


boot_straps_uw = 
  sleep_df|> 
  filter(university == "UW") |> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as.tibble),
    models = map(strap, \(df) lm(term_gpa ~ demo_race + demo_gender + demo_firstgen + bedtime_mssd + total_sleep_time + midpoint_sleep + daytime_sleep + cum_gpa, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)

boot_result_uw = 
  boot_straps_uw |> 
  group_by(term) |> 
  summarise(
    boot_est = mean(estimate),
    std.error = sd(estimate),
    conf.low = quantile(estimate, .025),
    conf.high = quantile(estimate, .975),
  ) |> 
  mutate(estimate = boot_est) |> 
  select(term, estimate, std.error, conf.low, conf.high) |> 
  mutate(term = reorder(term, estimate),
         university = 'UW')|> 
  filter(term != "(Intercept)")


uni_coef = rbind(boot_result_cmu,
                 boot_result_ndu,
                 boot_result_uw) |> 
  mutate(university = factor(university, levels = c("CMU", "NDU", "UW"))) 

uni_coef |>
  ggplot(aes(x = term, y = estimate, color = university)) +
  geom_point(position = position_dodge(width = 0.7)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.5, 
                position = position_dodge(width = 0.7)) +
  coord_flip() +
  labs(
    title = "Effect Sizes of Predictors on Cumulative GPA",
    x = "Predictors",
    y = "Coefficient Estimate"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

```

`cum_gpa` is significant in all the university and `total_sleep_time` is still significant in NDU and UW. The impact of other variables shift between universities. The detailed results are shown below:


#### CMU{.tabset}


```{r}
boot_result_cmu |> 
  mutate(
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-university) |> 
  arrange(-estimate) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

term_order <- boot_result_cmu |> arrange(-estimate) 
term_order = term_order$term
```



#### NDU{.tabset}


```{r}
boot_result_ndu |> 
  mutate(
    term = factor(term, levels = term_order),
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-university) |> 
  arrange(term) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

```


---

#### UW{.tabset}


```{r}
boot_result_uw |> 
  mutate(
    term = factor(term, levels = term_order),
    conf_int = paste0("[",round(conf.low,6),", ",round(conf.high,6), "]"),
    estimate = round(estimate,6),
    std.error = round(std.error,6)) |> 
  select(-conf.low, -conf.high) |> 
  select(-university) |> 
  arrange(term) |> 
  kable(format = "html") |> 
  kable_styling(c("hover", "condensed"), full_width = FALSE) |> 
  column_spec(1, bold = TRUE, width = "4.5cm") 

```

---

## 4. Conclusion

According to the linear regression model result based on the pooled data, we conclude that the `total_sleep_time` has the most important impact on `term_gpa` among all the four sleeping measures. It has significant positive coefficient in both the sleeping measure only model and the the characteristic-included model, which implies having more sleep can potentially improve students' academic performance. Additionally, `midpoint_sleep` is significant negative related to `term_gpa`, indicating that sleeping earlier is helpful to get better GPA if given the same total sleep time.

Besides, among the characteristic variables, `cum_gpa` is crucial to the `term_gpa`. And we also find the linear model shfits between universities, which means the impact of a specific predictors on the `term_gpa` varies from university to university.

However, the characteristic variable is usually fixed, students should focus on adjusting their sleeping patterns to improve their academic performance. Two useful suggestions we proposed based on our analysis are *sleep enough* and *sleep early*.


